---
title: "Italy CSEP models 2022"
output: html_notebook
---


```{r, echo=FALSE}
library(sp)
library(sf)
library(inlabru)
library(INLA)
library(future.apply)
library(ggplot2)
library(rgeos)
library(rgdal)
library(raster)
library(terra)

plan(multisession)

source("C:\\Users\\kirst\\Documents\\inlabru_notebooks\\inlabru_earthquake_notebooks\\Synthetic_inlabru\\ETAS_sampling\\spatial_ETAS_utils.R")
source("C:\\Users\\kirst\\Documents\\inlabru_notebooks\\inlabru_earthquake_notebooks\\Data_import\\StatSeisFuncs.R")
source("C:\\Users\\kirst\\Documents\\inlabru_notebooks\\inlabru_earthquake_notebooks\\Data_import\\inlabru_import_functions.R")

crs_wgs84 <- CRS(SRS_string='EPSG:4326')
crs_Ita <- CRS(SRS_string='EPSG:7794')
crs_Ita_km <- fm_crs_set_lengthunit(crs_Ita, "km")
```

```{r, warning=FALSE}
Horus_m4 <- read.csv("C:/Users/kirst/inlabru-independent/input/Horus_GKDec_1960_2020.csv")
Horus_m4 <- subset(Horus_m4, Horus_m4$dates < as.POSIXct("2011-01-01"))

colnames(Horus_m4) <- c("n", "dates", "x", "y", "mag", "event_id", "Cluster_Number", "cluster_ref", "ms_ref")
coordinates(Horus_m4) <- c("x", "y")
proj4string(Horus_m4) <- CRS(SRS_string='EPSG:4326')

#Horus_m4$sigMw = 0
## Get mag uncertainties from original horus catalogue
for (i in 1:length(Horus_m4)){
  loc <- which(Horus_m4@coords[i,1] == HORUS_Ita_Catalog$Lon & Horus_m4@coords[i,2] == HORUS_Ita_Catalog$Lat & Horus_m4$mag[i] == HORUS_Ita_Catalog$Mw)
  Horus_m4$sig.mag[i] <- HORUS_Ita_Catalog$sigMw[loc]
}

m4dec_bval = bval_distro(Horus_m4$mag, mbin=0.01, uncert="normal", sd=Horus_m4$sig.mag, iter=10000)

Horus_m4_dec <- spTransform(Horus_m4, crs_Ita_km)
#hm4 <- crop(Horus_m4_dec, It_coll_poly_km)

It_coll_poly_km <- spTransform(It_coll_poly, crs_Ita_km)
sf_poly_km <- st_as_sf(It_coll_poly_km)
mesh_poly <- as(st_buffer(sf_poly_km, 50), "Spatial")

hm4 <- crop(Horus_m4_dec, It_coll_poly_km)

mesh3 <- inla.mesh.2d(boundary=mesh_poly, max.edge = 30, crs=crs_Ita_km)
ggplot() + gg(mesh3) + gg(It_coll_poly_km) + gg(hm4)
```
Load covariates and covariate functions
```{r}
strainratedata <- read.delim("C:/Users/kirst/Documents/Italy_background/Italy background/strainratedata.txt")

xl <- 2
xu <- 22
yl <- 35
yu <- 52
SR <- subset(strainratedata, strainratedata$V1 > yl & strainratedata$V1 < yu & strainratedata$V2 > xl & strainratedata$V2 < xu)
SRGrid <- as.data.frame(cbind(SR$V2, SR$V1, log10(SR$strain_rate)))

## These values are ever so slightly off a regular grid (< .01 lat/lon, usually << 0.001)
#EPOSGrid[1:2] <- lapply(EPOSGrid[1:2], function(x) as.numeric(round(x, digits=2)))

srrast <- terra::rast(SRGrid, type="xyz")

test3 <- terra::focal(srrast, w=41, fun="mean", na.rm=TRUE, na.policy="only", extend = TRUE)
crs(test3) <- "epsg:4326"

focal_inf <- terra::as.data.frame(test3, xy=TRUE, na.rm = FALSE)
focal_inf$focal_mean_norm <- (focal_inf$focal_mean - min(focal_inf$focal_mean))/(max(focal_inf$focal_mean) - min(focal_inf$focal_mean))

test4 <- rast(focal_inf)

StrainRateFn_focal3 <- function(x,y) {
   # turn coordinates into SpatialPoints object:
   spp = SpatialPoints(data.frame(x=x,y=y))
   # attach the appropriate coordinate reference system (CRS)
   proj4string(spp) = CRS(proj4string(crs_Ita_km))
   spp = spTransform(spp, CRS(SRS_string = 'EPSG:4326'))
   ## BUT we want to use terra here so need a spatvector (!)
   spt <- vect(spp)
   crs(spt)="epsg:4326"
   # Extract values at spp coords, from our elev SpatialGridDataFrame
   v = terra::extract(test4, spt)
   v[is.na(v)] = -2 # NAs are a problem! Remove them
   return(v$focal_mean_norm)
 }


StrainRateFn_focal3(7200, 4500)
```
```{r}
DEM_merge <- terra::rast("C:/Users/kirst/Documents/Italy_test/Italy_DEM_merged.TIF")

DEM_min <- -29.16
DEM_max <- 4790.073

DEM_norm_fn <- function(x,y) {
  # turn coordinates into SpatialPoints object:
  spp = SpatialPoints(data.frame(x=x,y=y))
  # attach the appropriate coordinate reference system (CRS)
  proj4string(spp) = CRS(proj4string(crs_Ita_km))
  spp = spTransform(spp, crs_wgs84)
  spp_vect <- vect(spp)
  crs(spp_vect)="epsg:4326"
  spp_proj2 <- terra::project(spp_vect, crs(DEM_merge))
  # Extract values at spp coords, from our elev SpatialGridDataFrame
  v = terra::extract(DEM_merge, spp_proj2)
  v$Band_1[is.na(v$Band_1)] = 0 # NAs are a problem! Remove them
  val <- (v$Band_1 - DEM_min)/(DEM_max - DEM_min)
  return(val)
}

DEM_norm_fn(7200, 4500)

```
```{r}
Faults3 <- vect("H:/Users/Kirsty/Documents/italy-forecast/Data/DISS_3.2.1_shp/SHP/CSSPLN321.shp")
sf_faults <- st_as_sf(Faults3)

FaultDistFn= function(x, y) {
  # turn coordinates into SpatialPoints object
  ## Really need to figure out how to just make an sf object first!
  spp <- SpatialPoints(data.frame(x=x, y=y)) 
  sf_spp <- st_as_sf(spp)
  # set crs 
  st_crs(sf_spp) <- crs_Ita_km
  sf_spp <- st_transform(sf_spp, crs(Faults3))
  ## new fault distance calculation
  FD <- sf::st_distance(sf_faults, sf_spp)
  ## This gives us a distance to all fault polygons - we just want the closest one
  ## Also convert to km
  min_dist <- apply(FD, 2,  min)/1000

  return(min_dist)
}

## plot this as a sanity test
hist(FaultDistFn(hm4$x, hm4$y))
```

```{r}

FaultDistFn_lonlat= function(x, y) {
  # turn coordinates into SpatialPoints object
  ## Really need to figure out how to just make an sf object first!
  spp <- SpatialPoints(data.frame(x=x, y=y)) 
  sf_spp <- st_as_sf(spp)
  # set crs 
  st_crs(sf_spp) <- crs_wgs84
  ## new fault distance calculation
  FD <- sf::st_distance(sf_faults, sf_spp)
  ## This gives us a distance to all fault polygons - we just want the closest one
  ## Also convert to km
  min_dist <- apply(FD, 2,  min)/1000

  return(min_dist)
}

Test_grid <- expand.grid(x=seq(from=xl, to= xu, by=0.1), y=seq(from=yl, to=yu, by=0.1))

FD_test <- Test_grid

#FDFn <- Vectorize(FaultDistFn)
FD_test$FD <- FaultDistFn_lonlat(FD_test$x, FD_test$y)
coordinates(FD_test) <- c("x", "y")
gridded(FD_test) <- TRUE

ggplot() + gg(FD_test) + scale_fill_gradientn(colours = RColorBrewer::brewer.pal(9, "RdPu")) + gg(It_coll_poly)

FaultDistNormFn= function(x, y) {
  # turn coordinates into SpatialPoints object
  ## Really need to figure out how to just make an sf object first!
  spp <- SpatialPoints(data.frame(x=x, y=y)) 
  
  sf_spp <- st_as_sf(spp)
  # set crs 
  st_crs(sf_spp) <- crs_Ita_km
  sf_spp <- st_transform(sf_spp, crs_wgs84)
  
  ## new fault distance calculation
  FD <- sf::st_distance(sf_faults, sf_spp)
  ## This gives us a distance to all fault polygons - we just want the closest one
  ## Also convert to km
  min_dist <- (apply(FD, 2,  min)/1000)/max(FD_test$FD)

  return(min_dist)
}

max(FaultDistNormFn(hm4$x, hm4$y))
```
```{r}
Hseism<- readRDS("C:/Users/kirst/Documents/Italy_background/Italy background/CPTI_model_km.RDS")
#Hseism <- model_hist.pred


Hseism$med_norm <- (Hseism$median - min(Hseism$median))/(max(Hseism$median) - min(Hseism$median))
HS_Norm_Fn <-  function(x, y){
  # turn coordinates into SpatialPoints object:
  spp = SpatialPoints(data.frame(x=x,y=y))
  # attach the appropriate coordinate reference system (CRS)
  proj4string(spp) = CRS(proj4string(crs_Ita_km))
  #spp = spTransform(spp, proj4string(Faults3))
  # Extract values at spp coords, from our elev SpatialGridDataFrame
  v = over(spp,Hseism)
  hi <- v$med_norm
  hi[is.na(hi)]=0
  return(hi)
}

HS_Norm_Fn(7200, 4500)
```
```{r}
Faults3$SRMa_norm <- (Faults3$SRmax- min(Faults3$SRmax))/(max(Faults3$SRmax) - min(Faults3$SRmax))


MaxSlipFn_sp = function(x,y){
  spp = SpatialPoints(data.frame(x=x,y=y))
  proj4string(spp) = crs_Ita_km
  spp = spTransform(spp, crs_wgs84)
  vt = over(spp, as(Faults3, 'Spatial'))$SRMa_norm
  vt[is.na(vt)]=0
  #vt1 <- factor(vt)
  return(vt)
}

MaxSlipFn_sp(7200, 4500)
```

```{r, warning=FALSE}
spde.model= inla.spde2.pcmatern(mesh3, prior.sigma=c(5, 0.01), prior.range = c(50, 0.01))

#bru(..., options=list(strategy = "simplified.laplace"))

SRhistMaSlip.mdl= coordinates~ SR(main=StrainRateFn_focal3(x, y), model="linear") + hs(main=HS_Norm_Fn(x, y), model="linear")+ MaS(main=MaxSlipFn_sp(x, y), model="linear") + Smooth(main=coordinates, model=spde.model) +  Intercept(1)
SRhistMaSlip.fit = lgcp(SRhistMaSlip.mdl, hm4,  domain = list(coordinates = mesh3), 
                        samplers=It_coll_poly_km, 
                        options = list(E = 50, control.inla(strategy="laplace")))

start_time <- Sys.time()
histInv.mdl= coordinates~ IFD(main=FaultDistNormFn(x,y), model="linear") + hs(main=HS_Norm_Fn(x, y), model="linear") + Smooth(main=coordinates, model=spde.model) +  Intercept(1)
histInv.fit = lgcp(histInv.mdl, hm4,  domain = list(coordinates = mesh3), 
                   samplers=It_coll_poly_km, 
                   options = list(E = 50, control.inla(strategy="laplace")))
end_time <- Sys.time()
end_time - start_time


SRInv.mdl= coordinates~ IFD(main=FaultDistNormFn(x,y), model="linear") + SR(main=StrainRateFn_focal3(x, y), model="linear") + Smooth(main=coordinates, model=spde.model) +  Intercept(1)
SRInv.fit = lgcp(SRInv.mdl, hm4,  domain = list(coordinates = mesh3), 
                 samplers=It_coll_poly_km, 
                 options = list(E = 50, control.inla(strategy="laplace")))

SRMaIFD.mdl= coordinates~ SR(main=StrainRateFn_focal3(x, y), model="linear") + IFD(main=FaultDistNormFn(x, y), model="linear")+ MaS(main=MaxSlipFn_sp(x, y), model="linear") + Smooth(main=coordinates, model=spde.model) +  Intercept(1)
SRMaIFD.fit = lgcp(SRMaIFD.mdl, hm4,  domain = list(coordinates = mesh3), 
                   samplers=It_coll_poly_km, 
                   options = list(E = 50, control.inla(strategy="laplace")))

start_time <- Sys.time()
SRhistDEM.mdl = coordinates ~ DEM(main=DEM_norm_fn(x, y), model="linear") + SR(main=StrainRateFn_focal3(x, y), model="linear") + hs(main=HS_Norm_Fn(x, y), model="linear") + Smooth(main=coordinates, model=spde.model) +  Intercept(1)
SRhistDEM.fit = lgcp(SRhistDEM.mdl, hm4,  domain = list(coordinates = mesh3), 
                samplers=It_coll_poly_km, 
                options = list(E = 50, control.inla(strategy="laplace")))
end_time <- Sys.time()
end_time- start_time

start_time <- Sys.time()
histDEM.mdl = coordinates ~ DEM(main=DEM_norm_fn(x, y), model="linear")  + hs(main=HS_Norm_Fn(x, y), model="linear") + Smooth(main=coordinates, model=spde.model) +  Intercept(1)
histDEM.fit = lgcp(histDEM.mdl, hm4,  domain = list(coordinates = mesh3), 
                samplers=It_coll_poly_km, 
                options = list(E = 50, control.inla(strategy="laplace")))
end_time <- Sys.time()
end_time- start_time

start_time <- Sys.time()
InvDEMMa.mdl = coordinates ~ DEM(main=DEM_norm_fn(x, y), model="linear")  + IFD(main=FaultDistNormFn(x, y), model="linear")+ MaS(main=MaxSlipFn_sp(x, y), model="linear") + Smooth(main=coordinates, model=spde.model) +  Intercept(1)
InvDEMMa.fit = lgcp(InvDEMMa.mdl, hm4,  domain = list(coordinates = mesh3), 
                samplers=It_coll_poly_km, 
                options = list(E = 50, control.inla(strategy="laplace")))
end_time <- Sys.time()
end_time- start_time

```
```{r, warning=FALSE}

start_time <- Sys.time()
InvDEMMa.mdl = coordinates ~ DEM(main=DEM_norm_fn(x, y), model="linear")  + IFD(main=FaultDistNormFn(x, y), model="linear")+ MaS(main=MaxSlipFn_sp(x, y), model="linear") + Smooth(main=coordinates, model=spde.model) +  Intercept(1)
InvDEMMa.fit = lgcp(InvDEMMa.mdl, hm4,  domain = list(coordinates = mesh3), 
                samplers=It_coll_poly_km, 
                options = list(E = 50, control.inla(strategy="laplace")))
end_time <- Sys.time()
end_time- start_time

start_time <- Sys.time()
Pred_SRhistMa <- predict(SRhistMaSlip.fit, pixels(mesh3, mask=It_coll_poly_km), ~ (SR + hs + MaS + Smooth + Intercept))
end_time <- Sys.time()
end_time- start_time

start_time <- Sys.time()
Pred_histInv <- predict(histInv.fit, pixels(mesh3, mask=It_coll_poly_km), ~ (hs + IFD + Smooth + Intercept))
end_time <- Sys.time()
end_time- start_time

start_time <- Sys.time()
Pred_SRInv <- predict(SRInv.fit, pixels(mesh3, mask=It_coll_poly_km), ~ (SR + IFD + Smooth + Intercept))
end_time <- Sys.time()
end_time- start_time

start_time <- Sys.time()
Pred_SRMaIFD <-  predict(SRMaIFD.fit, pixels(mesh3, mask=It_coll_poly_km), ~ (SR + MaS + IFD + Smooth + Intercept))
end_time <- Sys.time()
end_time- start_time

start_time <- Sys.time()
Pred_SRhistDEM <- predict(SRhistDEM.fit, pixels(mesh3, mask=It_coll_poly_km), ~ (SR + hs + DEM + Smooth + Intercept))
end_time <- Sys.time()
end_time- start_time

start_time <- Sys.time()
Pred_histDEM <- predict(histDEM.fit, pixels(mesh3, mask=It_coll_poly_km), ~ (hs + DEM + Smooth + Intercept))
end_time <- Sys.time()
end_time- start_time


start_time <- Sys.time()
Pred_InvDEMMa <- predict(InvDEMMa.fit, pixels(mesh3, mask=It_coll_poly_km), ~ (IFD + DEM + MaS + Smooth + Intercept))
end_time <- Sys.time()
end_time- start_time

```
```{r}
pred_list <- c(Pred_SRhistMa, Pred_histInv, Pred_SRInv, Pred_SRMaIFD, Pred_SRhistDEM, Pred_histDEM, Pred_InvDEMMa)
pred_names <- c("Pred_SRhistMa", "Pred_histInv", "Pred_SRInv", "Pred_SRMaIFD", "Pred_SRhistDEM", "Pred_histDEM", "Pred_InvDEMMa")

comp_pairplots(pred_list = pred_list, pred_names = pred_names , "YlOrRd", It_coll_poly_km, var_lims=c(-1.5, 1.5))
```

```{r, warning=FALSE}
ips <- ipoints(It_coll_poly_km, mesh3)
Pred_Num_SRhistMa <- predict(SRhistMaSlip.fit, ips, ~ sum(weight * exp(SR + hs + MaS + Smooth + Intercept)))
SRhistMa.pr <- predict(SRhistMaSlip.fit, pixels(mesh3, mask=It_coll_poly_km), ~(SR + hs + MaS + Smooth + Intercept))

Pred_Num_histInv <- predict(histInv.fit, ips, ~ sum(weight * exp( hs + IFD + Smooth + Intercept)))
histInv.pr <- predict(histInv.fit, pixels(mesh3, mask=It_coll_poly_km), ~(hs + IFD + Smooth + Intercept))

Pred_Num_SRMaIFD <- predict(SRMaIFD.fit, ips, ~ sum(weight * exp(SR + MaS + IFD + Smooth + Intercept)))
SRMaIFD.pr <- predict(SRMaIFD.fit, pixels(mesh3, mask=It_coll_poly_km), ~(SR + MaS + IFD + Smooth + Intercept))

Pred_Num_SRInv <- predict(SRInv.fit, ips, ~ sum(weight * exp(SR + IFD + Smooth + Intercept)))
SRInv.pr <- predict(SRInv.fit, pixels(mesh3, mask=It_coll_poly_km), ~(SR + IFD + Smooth + Intercept))

library(RColorBrewer)
ggplot2::theme_set(ggplot2::theme_bw())
lprange <- range(SRhistMa.pr$q0.025, SRhistMa.pr$q0.975, SRMaIFD.pr$q0.025, SRMaIFD.pr$q0.975, SRInv.pr$q0.025, SRInv.pr$q0.975, histInv.pr$q0.025, histInv.pr$q0.975)
csc <- scale_fill_gradientn(colours = brewer.pal(9, "YlOrRd"), limits = lprange)

#europe <- ne_countries(continent = 'Europe', returnclass = "sf", scale = 'medium')


SRhistMaplt <- ggplot() + gg(SRhistMa.pr["q0.025"]) + ggtitle("SR + hist + maxslip") + csc + gg(It_coll_poly_km)
histInvplt <- ggplot() + gg(histInv.pr) + ggtitle("hist + IFD")  + csc + gg(It_coll_poly_km)
SRMaIFDplt <- ggplot() + gg(SRMaIFD.pr) + ggtitle("SR + maxslip + IFD") + csc + gg(It_coll_poly_km)
SRInvplt <- ggplot() + gg(SRInv.pr) + ggtitle("SR + IFD") + csc + gg(It_coll_poly_km)

multiplot(SRhistMaplt, histInvplt, SRMaIFDplt, SRInvplt, cols=2)
```

```{r, warning=FALSE}
samp_histInv <- generate(histInv.fit, mesh3, coordinates~(hs + IFD + Smooth + Intercept), n.samples = 10000)
start_time <- Sys.time()
samp_SRMaIFD <- generate(SRMaIFD.fit, mesh3, coordinates~(SR + MaS + IFD + Smooth + Intercept), n.samples = 10000)
end_time <- Sys.time()
end_time - start_time
start_time <- Sys.time()
samp_SRInv <- generate(SRInv.fit, mesh3, coordinates~(SR + IFD + Smooth + Intercept), n.samples = 10000)
end_time <- Sys.time()
end_time - start_time
## Not run yet...
start_time <- Sys.time()
samp_SRhistMA <- generate(SRhistMaSlip.fit, mesh3, coordinates~(SR + hs + MaS + Smooth + Intercept), n.samples = 10000)
end_time <- Sys.time()
end_time - start_time

start_time <- Sys.time()
Pred_Num_SRhistDEM <- predict(SRhistDEM.fit, ips, ~ sum(weight * exp(SR + hs + DEM + Smooth + Intercept)))
end_time <- Sys.time()
end_time - start_time

start_time <- Sys.time()
Pred_Num_InvDEMMa <- predict(InvDEMMa.fit, ips, ~ sum(weight * exp(IFD + DEM + MaS+ Smooth + Intercept)))
end_time <- Sys.time()
end_time - start_time

start_time <- Sys.time()
samp_SRhistDEM <- generate(SRhistDEM.fit, mesh3, coordinates~(SR + hs + DEM + Smooth + Intercept), n.samples = 10000)
end_time <- Sys.time()
end_time - start_time
start_time <- Sys.time()
samp_InvDEMMa <- generate(InvDEMMa.fit, mesh3, coordinates~(IFD + DEM + MaS+ Smooth + Intercept), n.samples = 10000)
end_time <- Sys.time()
end_time - start_time

```

```{r}
Horus_GKDec_1960_2020 <- read.csv("C:/Users/kirst/inlabru-independent/input/Horus_GKDec_1960_2020.csv")
Hset <- subset(Horus_GKDec_1960_2020, Horus_GKDec_1960_2020$dates < as.POSIXct("2011-01-01"))
m4dec_bval = bval_distro(Hset$mag, mbin=0.01, uncert="normal", sd=Hset$sig.mag, iter=10000)
hist(m4dec_bval$b_vals)
```


```{r, warning=FALSE}

samp_points_SRMAIFD_6010_1yr <- future_apply(samp_SRMaIFD, 2, point_sampler_best, bdy=It_coll_poly_km, mesh=mesh3, crs=crs_Ita_km, num_event=(Pred_Num_SRhistMa$mean), b_val=m4dec_bval$b_vals, m_min=4, corner_mag = 6.9, future.seed = 5)

Cats <- do.call(rbind, lapply(1:10000, function(x){
  data = spTransform(samp_points_SRMAIFD_6010_1yr[[x]], crs_wgs84)
  SpatialPointsDataFrame(coords = data@coords,
                         data = as.data.frame(data))
})
)

cat_lens <- as.numeric(unlist(lapply(samp_points_SRMAIFD_6010_1yr, length)))
Cats$cat_id <- rep(seq(1, length(samp_points_SRMAIFD_6010_1yr), by = 1), times=cat_lens)


times <- rep("2010-02-02T01:01:01.020000", length(Cats$mags))
depth <- rep(10, length(Cats$mags))
event_id <- seq(1, length(Cats$mags))
stoch_cat_set_SRMAIFD_6010_1yr <- as.data.frame(cbind(Cats$x, Cats$y, Cats$mags, times, depth, Cats$cat_id, event_id))

## Catalog readers are a bit fussy currently
write.table(format(stoch_cat_set_SRMAIFD_6010_1yr, digits = 6), "SRMAIFD_6010_1yr.dat", quote = FALSE, row.names = FALSE, col.names = FALSE, sep=",")

T210_Ml <- Mw_to_ML(as.numeric(stoch_cat_set_SRMAIFD_6010_1yr$V3))
stoch_cat_set_SRMAIFD_6010_1yr_Ml <- stoch_cat_set_SRMAIFD_6010_1yr
stoch_cat_set_SRMAIFD_6010_1yr_Ml$V3 <- T210_Ml
write.table(format(stoch_cat_set_SRMAIFD_6010_1yr_Ml, digits = 6), "SRMAIFD_6010_1yr_Ml.dat", quote = FALSE, row.names = FALSE, col.names = FALSE, sep=",")
```

```{r, warning=FALSE}
samp_points_histInv_6010_1yr <- future_apply(samp_histInv, 2, point_sampler_best, bdy=It_coll_poly_km, mesh=mesh3, crs=crs_Ita_km, num_event=(Pred_Num_histInv$mean), b_val=m4dec_bval$b_vals, m_min=4, corner_mag =6.9, future.seed = 5)

Cats <- do.call(rbind, lapply(1:10000, function(x){
  data = spTransform(samp_points_histInv_6010_1yr[[x]], crs_wgs84)
  SpatialPointsDataFrame(coords = data@coords,
                         data = as.data.frame(data))
})
)

cat_lens <- as.numeric(unlist(lapply(samp_points_histInv_6010_1yr, length)))
Cats$cat_id <- rep(seq(1, length(samp_points_histInv_6010_1yr), by = 1), times=cat_lens)


times <- rep("2010-02-02T01:01:01.020000", length(Cats$mags))
depth <- rep(10, length(Cats$mags))
event_id <- seq(1, length(Cats$mags))
stoch_cat_set_histInv_6010_1yr <- as.data.frame(cbind(Cats$x, Cats$y, Cats$mags, times, depth, Cats$cat_id, event_id))

## Catalog readers are a bit fussy currently
write.table(format(stoch_cat_set_histInv_6010_1yr, digits = 6), "histInv_6010_1yr.dat", quote = FALSE, row.names = FALSE, col.names = FALSE, sep=",")

histInv_Ml <- Mw_to_ML(as.numeric(stoch_cat_set_histInv_6010_1yr$V3))
stoch_cat_setT_histInv_6010_1yr_Ml <- stoch_cat_set_histInv_6010_1yr
#stoch_cat_set_histInv_6010_1yr_Ml$V3 <- histInv_Ml
write.table(format(histInv_Ml, digits = 6), "histInv_6010_1yr_Ml.dat", quote = FALSE, row.names = FALSE, col.names = FALSE, sep=",")
```

```{r, warning=FALSE}

samp_SRInv <- generate(SRInv.fit, mesh3, coordinates~(SR + IFD + Smooth + Intercept), n.samples = 10000)
end_time <- Sys.time()
end_time - start_time

samp_points_SRInv_6010_1yr <- future_apply(samp_SRInv, 2, point_sampler_best, bdy=It_coll_poly_km, mesh=mesh3, crs=crs_Ita_km, num_event=(Pred_Num_SRInv$mean), b_val=m4dec_bval$b_vals, m_min=4, corner_mag=6.9, future.seed = 5)

Cats <- do.call(rbind, lapply(1:10000, function(x){
  data = spTransform(samp_points_SRInv_6010_1yr[[x]], crs_wgs84)
  SpatialPointsDataFrame(coords = data@coords,
                         data = as.data.frame(data))
})
)

cat_lens <- as.numeric(unlist(lapply(samp_points_SRInv_6010_1yr, length)))
Cats$cat_id <- rep(seq(1, length(samp_points_SRInv_6010_1yr), by = 1), times=cat_lens)


times <- rep("2020-02-02T01:01:01.020000", length(Cats$mags))
depth <- rep(10, length(Cats$mags))
event_id <- seq(1, length(Cats$mags))
stoch_cat_set_SRInv_6010_1yr <- as.data.frame(cbind(Cats$x, Cats$y, Cats$mags, times, depth, Cats$cat_id, event_id))

## Catalog readers are a bit fussy currently
write.table(format(stoch_cat_set_SRInv_6010_1yr, digits = 6), "SRInv_6010_1yr_fixed.dat", quote = FALSE, row.names = FALSE, col.names = FALSE, sep=",")

SRInv_Ml <- Mw_to_ML(as.numeric(stoch_cat_set_SRInv_6010_1yr$V3))
stoch_cat_set_SRInv_6010_1yr_Ml <- stoch_cat_set_SRInv_6010_1yr
stoch_cat_set_SRInv_6010_1yr_Ml$V3 <- SRInv_Ml
write.table(format(stoch_cat_set_SRInv_6010_1yr_Ml, digits = 6), "SRInv_6010_1yr_Ml_fixed.dat", quote = FALSE, row.names = FALSE, col.names = FALSE, sep=",")
```

```{r, warning=FALSE}
samp_points_SRhistMA_6010_1yr <- future_apply(samp_SRhistMA, 2, point_sampler_best, bdy=It_coll_poly_km, mesh=mesh3, crs=crs_Ita_km, num_event=(Pred_Num_SRhistMa$mean), b_val=m4dec_bval$b_vals, m_min=4, corner_mag = 6.9, future.seed = 5)
###ru  from here

Cats <- do.call(rbind, lapply(1:10000, function(x){
  data = spTransform(samp_points_SRhistMA_6010_1yr[[x]], crs_wgs84)
  SpatialPointsDataFrame(coords = data@coords,
                         data = as.data.frame(data))
})
)

cat_lens <- as.numeric(unlist(lapply(samp_points_SRhistMA_6010_1yr, length)))
Cats$cat_id <- rep(seq(1, length(samp_points_SRhistMA_6010_1yr), by = 1), times=cat_lens)


times <- rep("2020-02-02T01:01:01.020000", length(Cats$mags))
depth <- rep(10, length(Cats$mags))
event_id <- seq(1, length(Cats$mags))
stoch_cat_set_SRhistMA_6010_1yr <- as.data.frame(cbind(Cats$x, Cats$y, Cats$mags, times, depth, Cats$cat_id, event_id))

## Catalog readers are a bit fussy currently
write.table(format(stoch_cat_set_SRhistMA_6010_1yr, digits = 6), "SRhistMA_6010_1yr.dat", quote = FALSE, row.names = FALSE, col.names = FALSE, sep=",")

SRhistMA_Ml <- Mw_to_ML(as.numeric(stoch_cat_set_SRMAIFD_6010_1yr$V3))
stoch_cat_setT_SRhistMA_6010_1yr_Ml <- stoch_cat_set_SRhistMA_6010_1yr
stoch_cat_setT_SRhistMA_6010_1yr_Ml$V3 <- SRhistMA_Ml
write.table(format(stoch_cat_setT_SRhistMA_6010_1yr_Ml, digits = 6), "SRhistMA_6010_1yr_Ml.dat", quote = FALSE, row.names = FALSE, col.names = FALSE, sep=",")
```

```{r, warning=FALSE}
samp_points_SRhistDEM_6010_1yr <- future_apply(samp_SRhistDEM, 2, point_sampler_best, bdy=It_coll_poly_km, mesh=mesh3, crs=crs_Ita_km, num_event=(Pred_Num_SRhistDEM$mean), b_val=m4dec_bval$b_vals, m_min=4, corner_mag = 6.9, future.seed = 5)
###ru  from here

Cats <- do.call(rbind, lapply(1:10000, function(x){
  data = spTransform(samp_points_SRhistDEM_6010_1yr[[x]], crs_wgs84)
  SpatialPointsDataFrame(coords = data@coords,
                         data = as.data.frame(data))
})
)

cat_lens <- as.numeric(unlist(lapply(samp_points_SRhistDEM_6010_1yr, length)))
Cats$cat_id <- rep(seq(1, length(samp_points_SRhistDEM_6010_1yr), by = 1), times=cat_lens)


times <- rep("2020-02-02T01:01:01.020000", length(Cats$mags))
depth <- rep(10, length(Cats$mags))
event_id <- seq(1, length(Cats$mags))
stoch_cat_set_SRhistDEM_6010_1yr <- as.data.frame(cbind(Cats$x, Cats$y, Cats$mags, times, depth, Cats$cat_id, event_id))

## Catalog readers are a bit fussy currently
write.table(format(stoch_cat_set_SRhistDEM_6010_1yr, digits = 6), "SRhistDEM_6010_1yr_v2.dat", quote = FALSE, row.names = FALSE, col.names = FALSE, sep=",")

SRhistDEM_Ml <- Mw_to_ML(as.numeric(stoch_cat_set_SRhistDEM_6010_1yr$V3))
stoch_cat_setT_SRhistDEM_6010_1yr_Ml <- stoch_cat_set_SRhistDEM_6010_1yr
stoch_cat_setT_SRhistDEM_6010_1yr_Ml$V3 <- SRhistDEM_Ml
write.table(format(stoch_cat_setT_SRhistDEM_6010_1yr_Ml, digits = 6), "SRhistDEM_6010_1yr_v2_Ml.dat", quote = FALSE, row.names = FALSE, col.names = FALSE, sep=",")
```


```{r, warning=FALSE}
samp_points_InvDEMMa_6010_1yr <- future_apply(samp_InvDEMMa, 2, point_sampler_best, bdy=It_coll_poly_km, mesh=mesh3, crs=crs_Ita_km, num_event=(Pred_Num_InvDEMMa$mean), b_val=m4dec_bval$b_vals, m_min=4, corner_mag = 6.9, future.seed = 5)
###ru  from here

Cats <- do.call(rbind, lapply(1:10000, function(x){
  data = spTransform(samp_points_InvDEMMa_6010_1yr[[x]], crs_wgs84)
  SpatialPointsDataFrame(coords = data@coords,
                         data = as.data.frame(data))
})
)

cat_lens <- as.numeric(unlist(lapply(samp_points_InvDEMMa_6010_1yr, length)))
Cats$cat_id <- rep(seq(1, length(samp_points_InvDEMMa_6010_1yr), by = 1), times=cat_lens)


times <- rep("2020-02-02T01:01:01.020000", length(Cats$mags))
depth <- rep(10, length(Cats$mags))
event_id <- seq(1, length(Cats$mags))
stoch_cat_set_InvDEMMa_6010_1yr <- as.data.frame(cbind(Cats$x, Cats$y, Cats$mags, times, depth, Cats$cat_id, event_id))

## Catalog readers are a bit fussy currently
write.table(format(stoch_cat_set_InvDEMMa_6010_1yr, digits = 6), "IvDMMa_6010_1yr.dat", quote = FALSE, row.names = FALSE, col.names = FALSE, sep=",")

InvDEMMa_Ml <- Mw_to_ML(as.numeric(stoch_cat_set_InvDEMMa_6010_1yr$V3))
stoch_cat_setT_InvDEMMa_6010_1yr_Ml <- stoch_cat_set_InvDEMMa_6010_1yr
stoch_cat_setT_InvDEMMa_6010_1yr_Ml$V3 <- InvDEMMa_Ml
write.table(format(stoch_cat_setT_InvDEMMa_6010_1yr_Ml, digits = 6), "IvDMMa_6010_1yr_Ml.dat", quote = FALSE, row.names = FALSE, col.names = FALSE, sep=",")
```

Add depths if we have them...
```{r}
coordinates(Cats) <- c("x", "y")
proj4string(Cats) <- CRS(SRS_string='EPSG:4326')

#Faults_sp <- readOGR("H:/Users/Kirsty/Documents/italy-forecast/Data/DISS_3.2.1_shp/SHP/CSSPLN321.shp")
#proj4string(Faults_sp) <- CRS(SRS_string='EPSG:4326')

## set to max depth
max_depth <- over(Cats, Faults_sp)$MaxDepth
max_depth[is.na(max_depth)] = 10

## set to min depth
min_depth <- over(Cats, Faults_sp)$MinDepth
min_depth[is.na(min_depth)] = 10

## Some random depth between the two
rand_depth <- runif(length(min_depth), min=min_depth, max = max_depth)

```


### Now let's think about where to put the largest events.
Options:
* Largest events go in highest intensity areas - this is trickier because you need to go back to your spatial intensity map and extract values (but I think this is not impossible...)
* Largest events are events which are on faults (if at all)

```{r}
## lets start with just one cat until we get the hang of it. 
## Does stacking MGR distros give you an MGR distro?
## Intuition says probably yes, but not necessarily the other way around!
testCat <- subset(stoch_cat_set_histInv_6010_1yr, stoch_cat_set_histInv_6010_1yr$V6 == 1 )
hist(as.numeric(testCat$V3))

testCat$V1 <- as.numeric(testCat$V1)
testCat$V2 <- as.numeric(testCat$V2)
coordinates(testCat) <- c("V1", "V2")
proj4string(testCat) <- CRS(SRS_string='EPSG:4326')

## Load fault info as sp
## To do: figure out a terra alt for this to avoid future disaster!
Faults_sp <- readOGR("H:/Users/Kirsty/Documents/italy-forecast/Data/DISS_3.2.1_shp/SHP/CSSPLN321.shp")
proj4string(Faults_sp) <- CRS(SRS_string='EPSG:4326')

## Return fault information for each event location
fault_inf <- over(testCat, Faults_sp)

## Order of max magnitudes related to specific faults
new_order <- sort(fault_inf$MaxMag, decreasing = TRUE, index.return=TRUE,na.last = TRUE)
## Order of magnitudes
mag_order <- sort(as.numeric(testCat$V3), decreasing = TRUE, index.return=TRUE)
#vector[order(index)]
## sort so largest magnitude is now on fault with highest max mag
new_mag_order <- mag_order$x[order(new_order$ix)]
```

Apply to whole cat object...
```{r}
fault_inf_cats <- over(Cats, Faults_sp)
new_order <- sort(fault_inf_cats$MaxMag, decreasing = TRUE, index.return=TRUE,na.last = TRUE)
## Order of magnitudes
mag_order <- sort(as.numeric(Cats$mags), decreasing = TRUE, index.return=TRUE)
#vector[order(index)]
## sort so largest magnitude is now on fault with highest max mag
new_mag_order <- mag_order$x[order(new_order$ix)]

plot(new_order$x, new_mag_order)
```
```{r}
plot(Cats$mags, new_mag_order)
```
```{r}
hist(Cats$mags - new_mag_order)
```
SO now largest events should be on faults with large max magnitude - none of our tests currently check this though...

```{r}
generate(histInv.fit, mesh3, coordinates~(hs + IFD + Smooth + Intercept), n.samples = 10000)

samp_InvDEMMa <- generate(InvDEMMa.fit, mesh3, coordinates~(IFD + DEM + MaS+ Smooth + Intercept), n.samples = 10000)


proj <- INLA::inla.mesh.project(mesh3, Cats)
lambda_ratio <- exp(as.vector(proj$A %*% loglambda) - loglambda_max)
```
```{r}
point_sampler_best <- function(loglambda, bdy, mesh,  crs, num_events, b_val, m_min, corner_mag){
  ## Number of events for single catalogue from a poisson distribution with lambda = num_events
  cat_n <- rpois(1, num_events)
  if (cat_n < 1){cat_n =1}
  ll <- as.data.frame(cbind(mesh$loc[,1], mesh$loc[,2], loglambda))
  colnames(ll) <- c("x", "y", "loglambda")
  coordinates(ll) <- c("x", "y")
  proj4string(ll) <- crs
  ## Mesh often extends outside our area of interest, crop so we are only considering sampling area bdy
  llcrop <- crop(ll, bdy)
  loglambda_max <- max(llcrop$loglambda)
  
  #print(loglambda_max)
  ## Set up a spatialpoints dataframe for our results
  samp.points <- SpatialPoints(data.frame(x = 0, y = 0))
  samp.points$mags <- 0
  samp.points <- samp.points[-1,]
  proj4string(samp.points) <- crs
  #print(proj4string(samp.points))
  num <- 0
  n1 <- 300000
  n2 <- 5000000
  ## To sample the correct number of points, keep going until the num >= cat_n
  while (num < cat_n){
    pts <- spsample(bdy, n1, "random")
    #pts <- spTransform(points, crs)
    proj <- INLA::inla.mesh.project(mesh, pts)
    lambda_ratio <- exp(as.vector(proj$A %*% loglambda) - loglambda_max)
    #print(max(lambda_ratio))
    #print(sum(proj$ok))
    keep <- proj$ok & (runif(n1) <= lambda_ratio)
    kept <- pts[keep]
    
    ## If we didn't get any events, run again with more sampled points
    while (length(kept) == 0){
      #print("No events kept - trying more locations")
      pts <- spsample(bdy, n2, "random")
      
      proj <- INLA::inla.mesh.project(mesh, pts)
      lambda_ratio <- exp(as.vector(proj$A %*% loglambda) - loglambda_max)
      #print(max(lambda_ratio))
      #print(sum(proj$ok))
      keep <- proj$ok & (runif(n2) <= lambda_ratio)
      kept <- pts[keep]
    }
    kept$mags <- rep(0, length(kept))
    
    samp.points <- rbind(samp.points, kept)
    num <- length(samp.points)
    
  }
  
  ## Keep exactly cat_n points, choose these randomly from all of the points we've kept so far
  kp <- sample(seq(1, length(samp.points), by=1), cat_n, replace=FALSE)
  samp.points <- samp.points[kp,]
  
  samp.points.proj <- INLA::inla.mesh.project(mesh, samp.points)
  ## Get magnitudes for this catalogue
  ## If b_val is a list of possible b-values, select one at random. If it's just one b-value, use that.
  b_est <- b_val[runif(1, 1, length(b_val))]
  
  mags <- TapGRM(cat_n, b_est, corner_mag, m_min)
  ## Now allocate magnitudes based on highest intensity, so highest magnitude is in area with highest intensity of kept catalogue
  
  return(samp.points)
}

```

```{r}
#future_apply(samp_InvDEMMa, 2, point_sampler_best, bdy=It_coll_poly_km, mesh=mesh3, crs=crs_Ita_km, num_event=(Pred_Num_InvDEMMa$mean), b_val=m4dec_bval$b_vals, m_min=4, corner_mag = 6.9, future.seed = 5)

#samp_InvDEMMa <- generate(InvDEMMa.fit, mesh3, coordinates~(IFD + DEM + MaS+ Smooth + Intercept), n.samples = 10000) 

ll <- as.data.frame(cbind(mesh3$loc[,1], mesh3$loc[,2], samp_InvDEMMa[,1]))
colnames(ll) <- c("x", "y", "loglambda")
coordinates(ll) <- c("x", "y")
proj4string(ll) <- crs_Ita_km
llcrop <- crop(ll, It_coll_poly_km)
loglambda_max <- max(llcrop$loglambda)

pts <- spsample(It_coll_poly_km, 10000, "random")
    #pts <- spTransform(points, crs)
proj <- INLA::inla.mesh.project(mesh3, pts)
lambda_ratio <- exp(as.vector(proj$A %*% samp_InvDEMMa[,1]) - max(samp_InvDEMMa[,1]))
    #print(max(lambda_ratio))
    #print(sum(proj$ok))
    keep <- proj$ok & (runif(n1) <= lambda_ratio)
    kept <- pts[keep]
    
    ## If we didn't get any events, run again with more sampled points
    while (length(kept) == 0){
      #print("No events kept - trying more locations")
      pts <- spsample(bdy, n2, "random")
      
      proj <- INLA::inla.mesh.project(mesh, pts)
      lambda_ratio <- exp(as.vector(proj$A %*% loglambda) - loglambda_max)
      #print(max(lambda_ratio))
      #print(sum(proj$ok))
      keep <- proj$ok & (runif(n2) <= lambda_ratio)
      kept <- pts[keep]
    }
    kept$mags <- rep(0, length(kept))
    
    samp.points <- rbind(samp.points, kept)
    num <- length(samp.points)
    

  
  ## Keep exactly cat_n points, choose these randomly from all of the points we've kept so far
  kp <- sample(seq(1, length(samp.points), by=1), cat_n, replace=FALSE)
  samp.points <- samp.points[kp,]
  
  samp.points.proj <- INLA::inla.mesh.project(mesh, samp.points)
  ## Get magnitudes for this catalogue
  ## If b_val is a list of possible b-values, select one at random. If it's just one b-value, use that.
  b_est <- b_val[runif(1, 1, length(b_val))]
  
  mags <- TapGRM(cat_n, b_est, corner_mag, m_min)
```
### Try using M25 seismicity
```{r}
Hseism<- readRDS("C:/Users/kirst/Documents/Italy_background/Italy background/CPTI_model_km.RDS")
#Hseism <- model_hist.pred


Hseism$med_norm <- (Hseism$median - min(Hseism$median))/(max(Hseism$median) - min(Hseism$median))
HS_Norm_Fn <-  function(x, y){
  # turn coordinates into SpatialPoints object:
  spp = SpatialPoints(data.frame(x=x,y=y))
  # attach the appropriate coordinate reference system (CRS)
  proj4string(spp) = CRS(proj4string(crs_Ita_km))
  #spp = spTransform(spp, proj4string(Faults3))
  # Extract values at spp coords, from our elev SpatialGridDataFrame
  v = over(spp,Hseism)
  hi <- v$med_norm
  hi[is.na(hi)]=0
  return(hi)
}

HS_Norm_Fn(7200, 4500)
```


